Practical Effects of Integrating Temperature with Strang Split Reactions

### 2021-02-22 strang/paper.tex
updates

\documentclass[times,modern]{aastex63}

@@

One option to make temperature integration less expensive is to freeze
the values of the thermodynamics quantities at the start of the
reactive integration.  This approach was done in \citet{ABNZ:III}.
This can work well in ``simmering'' phases of reactions where the
thermodynamic state is not changing much over a timestep, but for
explosive reactions, this may be a poor approximation.

@@

>>>>

\documentclass[times,preprint]{aastex63}

@@

The equations of hydrodynamics with reacting sources are:
\begin{align}
\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) &= 0 \\
\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho \Ub X_k) &= \rho \omegadot_k \\
\frac{\partial (\rho \Ub)}{\partial t} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &= 0 \\
\frac{\partial (\rho E)}{\partial t} + \nabla \cdot \left [ (\rho E + p) \Ub \right ] &= \rho \Sdot
\end{align}

When we are reacting, we need the temperature, so we can look at a temperature evolution
equation:
\begin{equation}
\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - e_\rho \right ) \frac{D\rho}{Dt} + \rho \Sdot
\end{equation}
(this form neglects the change in temperature due to composition
changes from reactions, see \citealt{ABNZ:III}, which assumes chemical
equilibrium).  During reactions, we neglect the hydrodynamics terms,
so the reactive system updates according to:
\begin{align}
\frac{D\rho}{Dt} &= 0 \\
\frac{DX_k}{Dt} &= \omegadot_k \label{eq:species} \\
\frac{DT}{Dt} &= \frac{\Sdot}{c_v}
\end{align}

There are several different approximations that people make in the
literature to this operator-split reacting system, in order to make
integrating the reaction system computationally less expensive:
\begin{itemize}
\item Neglect temerature evolution completely in the burn, only
  evolving Eq.~\ref{eq:species}.  This is the method used in
  \citet{flash}.
\item Evolve the temperature equation, but ``freeze'' the specific
  heat, $c_v$, at the start of integration to avoid potentially
  expensive equation of state calls.  This was discussed in
  \citet{ABNZ:III} and is the default method in \castro.
\item Evolve the full system, calling the equation of state each time
  we evaluate the righthand side of the system to get an updated
  $c_v$.
\end{itemize}
Depending on how vigorous the burning is and how much the temperature
changes during a hydrodynamic timestep, one or more of these
approximations may be reasonable.  For explosive reactions, we expect
that evolving the full system will be required.  The goal of this
note is to try to quantify the convergence of a reacting hydrodynamics
code with these different approximations.

@@

Looking at global convergence of a reacting hydrodynamics problem we
see that second order convergence is only realized when temperature is
evolved alongside the composition when using an operator-split
approach to reactions.
